Specific taxa differentiate proximal and distal microbiota in the unprepped

### 2017-07-05 submission/manuscript.Rmd
progress with restructuring intro

Several recent findings have shown that development and progression of CRC can be attributed to specific molecular events as a result of interactions between the gut microbiota and human host (@Yamauchi2012). Our group and others have found that the stool microbiome of patients with CRC is distinct from that of healthy people (@Zackular2014). Comparison of the specific bacteria present on CRC tumors with those found on nearby healthy tissue has also identified specific bacterial species that are tumor-associated (@Dejea2014). These species include the oral pathogens _Fusobacterium nucleatum_ and _Porphyromonas asacharolytica_. Interestingly, these periodontal pathogens have been highly predictive of whether a patient had CRC tumors or not in our classification studies (@Baxter2016).  Because these studies were performed in mice or examined only shed human stool, they were unable to analyze paried samples from the proximal and distal sides of the colon. Similarly, comparisons of on- or off-tumor bacteria rarely have matched tissue from the other side of the colon from the same patient, limiting what conclusions can be drawn about the colonic microbiome overall, let alone at that specific site. Due to these limitations, the contribution of the gut microbiota to these subtypes is largely undefined. Characterizing these communities could provide needed insight into CRC etiology, including how the disruption of the healthy community could promote the initiation or proliferation of the distinct distal and proximal CRC tumors.

@@

why are L/R M/L interesting questions? because diseases progress differently depending on where they are located.

>>>>

Several recent findings have shown that development and progression of IBD or CRC can be attributed to specific molecular events as a result of interactions between the gut microbiota and human host (@Yamauchi2012, IBDREF). For instance, comparison of the bacteria present on CRC tumors with those found on nearby healthy tissue has identified specific species that are tumor-associated (@Dejea2014). These species include the oral pathogens _Fusobacterium nucleatum_ and _Porphyromonas asacharolytica_. Interestingly, these periodontal pathogens have been highly predictive of whether a patient had CRC tumors or not in our human stool classification studies (@Baxter2016). _F. nucleatum_ has also been found to be elevated in the stool and biopsies of patients with IBD as compared to healthy controls (Strauss2011). Furthermore, studies of _F. nucleatum_ isolated from mucosal biopsies showed that more invasive  _F. nucleatum_ positively correlates with IBD disease level (Strauss2011). The bacteria appear to have a high-impact despite being lowly-abundant-. Thus the molecular capabilities of these rare taxa may drive the colonic disease state. Previous studies examining the microbes associated with CRC and IBD have focused on overall community structure, diversity and membership and not lowly-abundant species. Further, because these studies often examined only shed human stool or one biopsy at a time, they were unable to analyze paried samples from the proximal and distal sides of the colon. Similarly, comparisons of on- or off-tumor bacteria rarely have matched tissue from the other side of the colon from the same patient, limiting what conclusions can be drawn about the colonic microbiome overall, let alone at that specific site. Due to these limitations, the contribution of the gut microbiota to these subtypes is largely undefined. Characterizing these communities could provide needed insight into CRC etiology, including how the disruption of the healthy community could promote the initiation or proliferation of the distinct distal and proximal CRC tumors.

@@
